---
title: "Week 4 homework solutions"
author: "Hanusia Higgins"
date: "2/5/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Suppose x = 1.1, a = 2.2, and b = 3.3. Assign each expression to the value of the variable zand print the value stored in z.  
a. $x^{a^b}$  
b. $(x^a)^b$  
c. $3x^2+2x^2+1$  


```{r}
x <- 1.1 #assigning variable values as directed
a <- 2.2
b <- 3.3
z <- x^a^b #assigning value of first expression to z
print(z) #printing the result
z <- (x^a)^b #repeating for the second expression
print(z)
z <- 3*x^2+2*x^2+1 #repeating for the third expression
print(z)
```


2. Using the `rep` and `seq` functions, create the following vectors:  
a. $(1,2,3,4,5,6,7,8,7,6,5,4,3,2,1)$  
b. $(1,2,2,3,3,3,4,4,4,4,5,5,5,5,5)$  
c. $(5,4,4,3,3,3,2,2,2,2,1,1,1,1,1)$  

```{r}
a <- c(seq(1,8), seq(7,1)) #create first vector of ascending and descending sequences
print(a)
b <- rep(seq(1,5), times=seq(1,5)) #create second vector of sequence with each number n repeated n times
print(b)
c <- rep(seq(5,1), times=seq(1,5)) #create third vector of sequence with each number repeated the number of times of the sequence in reverse
print(c)
```

3. Create a vector of two random uniform numbers. In a spatial map, these can be interpreted as x and y coordinates that give the location of an individual (such as a marked forest tree in a plot that has been mapped). Using one of R’s inverse trigonometry functions `(asin(), acos()`, or `atan()`), convert these numbers into polar coordinates (If you don’t know what polar coordinates are, read about them on the web or in your calculus textbook).
```{r}
z <- runif(2) #creating a vector of two random uniform numbers, representing x and y
print(paste("Cartesian coordinates: (",z[1], ", ", z[2], ")"), quote=FALSE)
r <- sqrt(z[1]^2+z[2]^2) #converting to the value of r, the hypotenuse of the triangle formed by x and y
theta <- atan(z[2]/z[1]) #converting to the value of theta, the inverse tangent of y/x
print(paste("polar coordinates: (",r, ", ", theta, ")"), quote=FALSE)
```

4. Suppose that `queue <- c("sheep", "fox", "owl", "ant")` and that queue represents the animals that are lined up to enter Noah’s Ark, with the sheep at the front of the line. Using R expressions, update the queue successively as  
  a. the serpent arrives;  
  b. the sheep enters the ark;  
  c. the donkey arrives and talks his way to the front of the line;  
  d. the serpent gets impatient and leaves;  
  e. the owl gets bored and leaves;  
  f. the aphid arrives and the ant invites him to cut in line.  
  g. Finally, determine the position of the aphid in the line.  

```{r}
queue <- c("sheep", "fox", "owl", "ant") #setting up the initial queue
print(queue)
queue <- c(queue,"snake") #the serpent arrives
print(queue)
queue <- queue[2:length(queue)] #the sheep enters the ark
print(queue)
queue <- c("donkey",queue) #the donkey arrives and talks his way to the front of the line
print(queue)
queue <- queue[1:length(queue)-1] #the serpent gets impatient and leaves
print(queue)
queue <- queue[c(1,2,4)] #the owl gets bored and leaves
print(queue)
queue <- c(queue[1:2],"aphid",queue[3]) #the aphid arrives and the ant invites him to cut in line.
print(queue)
which(queue=="aphid") #Finally, determine the position of the aphid in the line
```

5. Use R to create a vector of all of the integers from 1 to 100 that are not divisible by 2, 3, or 7.

```{r}
z <- seq(1:100) #starting with the integers from 1 to 100
z <- z[z%%2!=0] #filtering out the ones that are divisible by 2
z <- z[z%%3!=0] #filtering out the ones that are divisible by 3
z <- z[z%%7!=0] #filtering out the ones that are divisible by 7
print(z) #printing the final vector
```





