---
title: "Homework 7 solutions"
author: "Hanusia Higgins"
date: "2/26/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Generating the false data set

This is a simplified version of how I imagine some of my data will look.  
(I haven't actually collected the data yet)  

The y-variable is the number of ash tree (Fraxinus sp.) seedlings are collected in a sample plot.  

The x-variable I will look at today is:   

Harvest status (harvested for emerald ash borer management/not harvested)     
Which will be represented as the following four groups:  

1. no harvest: "control"   
2. has been harvested: "harvest"  

```{r}
#following the same format we learned in class...
n_group <- 2 #number of treatment groups
n_name <- c("control","harvest") #naming the treatment groups
n_size <- c(25,25) #sample size = 25 of each treatment type

#I predict a higher # of seedlings in control plots than harvested ones
n_mean <- c(45,30)
n_sd <- c(10,10) #same SD/variance between groups
t_group <- rep(n_name,n_size)

id <- 1:(sum(n_size))

#simulating data as the response variable:
seedlings <- c(rnorm(n=n_size[1],mean=n_mean[1],sd=n_sd[1]),
             rnorm(n=n_size[2],mean=n_mean[2],sd=n_sd[2])) 

#creating the data frame
sim_data <- data.frame(id,t_group,seedlings)
str(sim_data)

```

### Analyzing the data

I will use an ANOVA to analyze the data.  
(I could also use a t-test since there are only 2 groups, but I may want to adjust this to include more treatment groups in the future.) 

```{r}
ano_model <- aov(seedlings~t_group,data=sim_data)
print(summary(ano_model)) #found a significant result
```

### Visualizing the data

I will use the ggplot2 package to visualize the data.

```{r}
library(ggplot2)

ano_plot <- ggplot(data=sim_data, #plotting the anova results
                   aes(x=t_group, #by treatment group
                       y=seedlings,
                   fill=t_group)) +
                   geom_boxplot() #specifies the type of graph we want
print(ano_plot)
```

### Varying the mean

In this section, I will vary the means of both groups to find out how close they can be while still showing a statistically significant difference. 
```{r}
#Keeping all else the same, besides the mean......

for (i in 0:7){

n_mean2 <- c(45,30)

  n_mean2[1] = n_mean2[1]-i #bringing the means closer together
  n_mean2[2] = n_mean2[2]+i #with each iteration of the for loop
  
  #the new means are reflected in the response variable distributions
  seedlings <- c(rnorm(n=n_size[1],mean=n_mean2[1],sd=n_sd[1]),
             rnorm(n=n_size[2],mean=n_mean2[2],sd=n_sd[2])) 

#creating the data frame
sim_data <- data.frame(id,t_group,seedlings)

#analyzing the data
ano_model <- aov(seedlings~t_group,data=sim_data)

#using code from last week as a model:
z <- summary(ano_model)

#for each iteration of effect size, this prints out the difference between the means, the associated p-value, and whether it's significant
print(paste("Effect size: ",
            n_mean[1]-n_mean[2],
            ", p-value: ",
            unlist(z)[9],
            ", significant? ",
            unlist(z)[9]<0.05))
}
```

##### After running this code several times, I found that the significance was variable at all but the highest and lowest effect sizes I chose.  
##### When the difference between the means was 15, the difference in the distributions was always significant. 
##### When the difference between the means was 1, the difference in the distributions was never significant.
##### For all the effect sizes in between, it varied...
I could also add an outer for loop to run this code a bunch of times and find out how likely each effect size is to be significant.

### Varying the sample size

In this section, I will vary the sample size to find out how small it can be while still showing a statistically significant difference between the treatment groups (using my original effect size).

```{r}
#Keeping all else the same, besides the sample size...

for (i in seq(from=25,to=5,by=-5)){
  
n_size <- c(i,i)

t_group <- rep(n_name,n_size)
id <- 1:(sum(n_size))

#simulating data as the response variable:
seedlings <- c(rnorm(n=n_size[1],mean=n_mean[1],sd=n_sd[1]),
             rnorm(n=n_size[2],mean=n_mean[2],sd=n_sd[2])) 

#creating the data frame
sim_data <- data.frame(id,t_group,seedlings)

#analyzing the data
ano_model <- aov(seedlings~t_group,data=sim_data)

#using code from last week as a model:
z <- summary(ano_model)

#for each iteration of effect size, this prints out the difference between the means, the associated p-value, and whether it's significant
print(paste("Sample size per group: ",
            i,
            ", p-value: ",
            unlist(z)[9],
            ", significant? ",
            unlist(z)[9]<0.05))
}
```

##### When the sample size per group was 10 or greater, the difference in the distributions was always significant. 
##### However, at a sample size per group of 5, the difference in the distributions was sometimes significant, other times not.
##### Therefore, the minimum sample size I would need to detect a statistically significant effect is 10 samples per treatment group.


