---
title: "Week 8 homework solutions"
author: "Hanusia Higgins"
date: "3/4/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Data manipulations using the `dplyr` package

This homework assignment focus on data manipulation in `R`. Complete these problems using the `dplyr` package, and if you have time, try each of these exercises using the subsetting methods in base `R` that have been covered in lecture.  
  
1. Examine the structure of the `iris` dataset. How many observations and variables are in the dataset?
```{r}
library(dplyr) #loading in dplyr package
glimpse(iris) #taking an initial look at this dataset
#based on the glimpse() results:
#this dataset has 150 observations of 5 variables
```

2. Create a new data frame `iris1` that contains only the species virginica and versicolor with sepal lengths longer than 6 cm and sepal widths longer than 2.5 cm. How many observations and variables are in the dataset?
```{r}
iris1 <- filter(iris, Species==c("virginica", "versicolor"), Sepal.Length>6, Sepal.Width>2.5) #filtering iris dataset based on the given parameters; in this case, commas stand in for "&"
glimpse(iris1) #checking out the new dataset
#iris1 has 28 observations of 5 variables
```

3. Now, create a `iris2` data frame from `iris1` that contains only the columns for Species, Sepal.Length, and Sepal.Width. How many observations and variables are in the dataset?
```{r}
#selecting the desired columns
iris2 <- select(iris1, Species, Sepal.Length, Sepal.Width)
glimpse(iris2) #checking out the new dataset
#iris2 has 28 observations of 3 variables
```

4. Create an `iris3` data frame from `iris2` that orders the observations from largest to smallest sepal length. Show the first 6 rows of this dataset.
```{r}
iris3 <- arrange(iris2, by=desc(Sepal.Length)) #reordering observations in descending order of sepal length
head(iris3, 6) #viewing the first 6 rows of the dataset
```

5. Create an `iris4` data frame from `iris3` that creates a column with a sepal area (length * width) value for each observation. How many observations and variables are in the dataset?
```{r}
iris4 <- mutate(iris3, Sepal.Area=Sepal.Length*Sepal.Width)
#adding a column of sepal area (length*width) for each obs
glimpse(iris4)
#iris4 has 28 observations of 4 variables
```

6. Create `iris5` that calculates the average sepal length, the average sepal width, and the sample size of the entire `iris4` data frame and print `iris5`.
```{r}
iris5 <- summarize(iris4, #creating iris5
                   #calculating avg sepal length:
                   mean_sepal_length = mean(Sepal.Length), 
                   #calculating avg sepal width:
                   mean_sepal_width = mean(Sepal.Width),
                   #calculating sample size with the helper function n():
                   sample_size = n())
print(iris5)
```

7. Finally, create `iris6` that calculates the average sepal length, the average sepal width, and the sample size for each species of in the `iris4` data frame and print `iris6`.
```{r}
iris6 <- group_by(iris4, Species) #grouping by species
iris6 <- summarize(iris6, #now summarizing that by species
                   #calculating avg sepal length:
                   mean_sepal_length = mean(Sepal.Length), 
                   #calculating avg sepal width:
                   mean_sepal_width = mean(Sepal.Width),
                   #calculating sample size with the helper function n():
                   sample_size = n())
print(iris6)
```

8. In these exercises, you have successively modified different versions of the data frame `iris1 iris1 iris3 iris4 iris5 iris6`. At each stage, the output data frame from one operation serves as the input fro the next.

A more efficient way to do this is to use the pipe operator %>% from the tidyr package. See if you can rework all of your previous statements into an extended piping operation that uses `iris` as the input and generates `iris6` as the output.
```{r}
#doing the same as all the above, via piping!
iris6 <- iris %>% 
  #refining initial dataset by species, sepal L & W
  filter(Species==c("virginica", "versicolor"), Sepal.Length>6, Sepal.Width>2.5) %>% 
  #selecting desired columns
  select(Species, Sepal.Length, Sepal.Width) %>% 
  #reordering observations in descending order of sepal length
  arrange(by=desc(Sepal.Length)) %>%  
  #adding a column of sepal area (length*width) for each obs
  mutate(Sepal.Area=Sepal.Length*Sepal.Width) %>% 
  group_by(Species) %>% 
  #creating this summary of the data so far, still grouped by species
  summarize(#calculating avg sepal length:
            mean_sepal_length = mean(Sepal.Length), 
            #calculating avg sepal width:
            mean_sepal_width = mean(Sepal.Width),
            #calculating sample size with the helper function n():
            sample_size = n()) 
  print(iris6)
```

