---
title: "Week 6 homework"
author: "Hanusia Higgins"
date: "2/19/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Open libraries
```{r}
library(ggplot2) # for graphics
library(MASS) # for maximum likelihood estimation
```

### Read in data vector
```{r}
# quick and dirty, a truncated normal distribution to work on the solution set

# z <- rnorm(n=3000,mean=0.2)
# z <- data.frame(1:3000,z)
# names(z) <- list("ID","myVar")
# z <- z[z$myVar>0,]
# str(z)
# summary(z$myVar)

#adding in real data
#from EDI data portal, "Future of Oak Forests experiment-
#DBH of trees in Black Rock Forest, New York, USA, 2007."
#by Schuster, W.S.

z <- read.table("BRf_NorthSlopeOaks2007.csv",header=TRUE,sep=",", stringsAsFactors=FALSE)
str(z) #it's counting the DBH values (continuous) as characters
summary(z) #so I need to get it to read as numeric

head(z)
#making the column I want a numeric vector
z$DBH <- as.numeric(z$DBH)  
#removing NA values
z <- na.omit(z)
str(z) #that's better!
#reducing it to 2 columns for simplicity
z <- z[,3:4]
#renaming the columns to fit in w/ the rest of the code
names(z) <- list("Species","myVar")
str(z)
summary(z)

```

### Plot histogram of data
```{r}
#visualizing the data before working with it
p1 <- ggplot(data=z, aes(x=myVar, y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) 
print(p1)
```

### Add empirical density curve
```{r}
#adding a density curve that follows the actual data
p1 <-  p1 +  geom_density(linetype="dotted",size=0.75)
print(p1)
```

### Get maximum likelihood parameters for `normal`
```{r}
#if these data do fit a normal distribution,
#what are the mean and variance?
normPars <- fitdistr(z$myVar,"normal")
print(normPars)
str(normPars)
normPars$estimate["mean"] # note structure of getting a named attribute
normPars$estimate["sd"] #doing the same thing for sd
```

### Plot `normal` probability density
```{r}
#this one doesn't fit very well.
meanML <- normPars$estimate["mean"]
sdML <- normPars$estimate["sd"]

xval <- seq(0,max(z$myVar),len=length(z$myVar))

 stat <- stat_function(aes(x = xval, y = ..y..), fun = dnorm, colour="red", n = length(z$myVar), args = list(mean = meanML, sd = sdML))
 p1 + stat
```

### Plot `exponential` probability density
```{r}
#this also doesn't fit very well.
#fitting the DBH data to an exponential distribution
expoPars <- fitdistr(z$myVar,"exponential")
#storing the exponential growth rate as a variable
rateML <- expoPars$estimate["rate"]

#graphin' it
stat2 <- stat_function(aes(x = xval, y = ..y..), fun = dexp, colour="blue", n = length(z$myVar), args = list(rate=rateML))
 p1 + stat + stat2
```

### Plot `uniform` probability density
```{r}
#this one doesn't fit well, either...the data definitely isn't uniform.
stat3 <- stat_function(aes(x = xval, y = ..y..), fun = dunif, colour="darkgreen", n = length(z$myVar), args = list(min=min(z$myVar), max=max(z$myVar)))
 p1 + stat + stat2 + stat3
```

### Plot `gamma` probability density
```{r}
#this distribution still doesn't capture the bimodal distribution.
#trying to fit to a gamma distribution
gammaPars <- fitdistr(z$myVar,"gamma")
#recording the gamma-specific variables of shape and rate
shapeML <- gammaPars$estimate["shape"]
rateML <- gammaPars$estimate["rate"]

stat4 <- stat_function(aes(x = xval, y = ..y..), fun = dgamma, colour="brown", n = length(z$myVar), args = list(shape=shapeML, rate=rateML))
 p1 + stat + stat2 + stat3 + stat4
```

### Plot `beta` probability density
```{r}
#first, transforming the data so it is all in the range of 0:1 (required for the beta distribution.)
pSpecial <- ggplot(data=z, aes(x=myVar/(max(myVar + 0.1)), y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) + 
  xlim(c(0,1)) +
  geom_density(size=0.75,linetype="dotted")

#now, fitting a beta distribution. This one is probably the best fit, even though it's still not capturing the two peaks of the data.
betaPars <- fitdistr(x=z$myVar/max(z$myVar + 0.1),start=list(shape1=1,shape2=2),"beta")
#recording the parameters of this distribution of best fit, which are called shape1 and shape2
shape1ML <- betaPars$estimate["shape1"]
shape2ML <- betaPars$estimate["shape2"]

#graphing it
statSpecial <- stat_function(aes(x = xval, y = ..y..), fun = dbeta, colour="orchid", n = length(z$myVar), args = list(shape1=shape1ML,shape2=shape2ML))
pSpecial + statSpecial
```

## Step 5.

I am using the beta distribution because it fits the data best (although it still doesn't really fit).It still doesn't capture the bimodal nature of these data.
```{r}
#getting the maximum likelihood parameters for the beta distribution: 
print(shape1ML)
print(shape2ML)

length(z$myVar) #getting length of the real data

#simulating a new data set using these parameters:
z2 <- rbeta(n=length(z$myVar), shape1=shape1ML, shape2=shape2ML)
z2 <- data.frame(1:length(z$myVar),z2)
names(z2) <- list("ID","myVar")
str(z2)
summary(z2$myVar)

#plotting the new simulated dataset in a histogram:
p1 <- ggplot(data=z2, aes(x=myVar, y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) 

#re-assigning xval to relate to the simulated data
xval <- seq(0,max(z2$myVar),len=length(z2$myVar))

#adding the probability density curve:
statSpecial <- stat_function(aes(x = xval, y = ..y..), fun = dbeta, colour="orchid", n = length(z2$myVar), args = list(shape1=shape1ML,shape2=shape2ML))

p1 + statSpecial #graphing simulated data w/ the curve

#now, plotting the original (real) data again (modified to fit into the beta distribution)
pSpecial <- ggplot(data=z, aes(x=myVar/(max(myVar + 0.1)), y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) +
  xlim(c(0,1))

#re-assigning xval to relate to the real data
xval <- seq(0,max(z$myVar),len=length(z$myVar))

#adding the probability density curve:
statSpecial <- stat_function(aes(x = xval, y = ..y..), fun = dbeta, colour="orchid", n = length(z$myVar), args = list(shape1=shape1ML,shape2=shape2ML))

pSpecial + statSpecial #graphing real data w/ the curve

```


The simulated data fits the beta distribution curve much better than the real data.  
The beta distribution still doesn't capture the bimodality of the real data, so by nature it is not doing a good job of simulating realistic data that match the original measurements.

